### 2013-11-29 NeurounitFrontiers/Paper.tex
Cyrus' final intro text.

\documentclass{frontiersSCNS}
%\documentclass{article}
%\usepackage{biblatex}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{times}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{listings}
%\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{setspace}
\usepackage{python_syntax}
%\usepackage{bibentry}
\usepackage{todonotes}
\usepackage{url}
\usepackage{lineno}
%\usepackage{natbib}
%\usepackage{caption}

%\usepackage{biblatex}
%\DeclareCiteCommand{\wcite}
%    [\mkbibfootnote]
%    {}
%    {\printfield{title}--\printfield{howpublished}}
%    {\addcomma\addspace}
%    {}

\usepackage[english]{babel} 
\def\UrlFont{\em} % Italicize all URLs.  
\lstset{
	basicstyle=\small\ttfamily,
	breaklines=true,
	language=Python,
	}
\let\verbx\lstinline
\newcommand{\bfhead}[1]{\noindent \textbf{#1:}}

\linenumbers
\copyrightyear{}
\pubyear{}

\def\journal{NeuroInformatics}%%% write here for which journal %%%
\def\DOI{}
\def\articleType{Methods}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Sample {et~al.}} %use et al only if is more than 1 author
\def\Authors{Richard C. Gerkin\,$^{1,*}$ and Cyrus Omar\,$^{2}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$School of Life Sciences, Arizona State University, Tempe, AZ, USA \\
$^{2}$Dept. of Computer Science, Carnegie Mellon University, Pittsburgh, PA, USA  }
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Rick Gerkin}
\def\corrAddress{ISTB-1 476}
\def\corrEmail{rgerkin@asu.edu}

% \color{FrontiersColor} Is the color used in the Journal name, in the title, and the names of the sections.

\begin{document}

\onecolumn
\firstpage{1}

\title[Running Title]{NeuronUnit: Collaboratively Testing the Validity of Neuroscientific Models}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
\topic{Python in Neuroscience II}% If your article is part of a Research Topic, please indicate here which.

\maketitle

\begin{abstract}
Rigorously validating a quantitative scientific model requires comparing its predictions against an unbiased selection of experimental observations according to sound statistical criteria. Developing new models thus requires a comprehensive and contemporary understanding of competing models, relevant data and statistical best practices. Today, developing such an understanding requires an encyclopedic knowledge of the literature. 
Unfortunately, in rapidly-growing fields like neuroscience, this is becoming increasingly untenable, even for the most conscientious scientists. For new scientists, this can be a significant barrier to entry.

%However, today's model validation process is based on one-time peer review of a scientific publication. , and thus remains an informal and incomplete process, particularly in fields like neuroscience where the number of competing models and the volume of relevant data is growing rapidly.   
%This makes it difficult to determine community-accepted validation criteria, compare competing models,  determine the state-of-the-art given the latest experimental data, and to precisely identify open modeling problems.

Software engineers seeking to verify, validate and contribute to a complex software project rely not only on volumes of human documentation, but on suites of simple executable tests, called ``unit tests''. 
Drawing inspiration from this practice, we have developed \textit{SciUnit}, an easy-to-use framework for developing ``model validation tests'' -- 
executable functions, here written in Python, that generate and statistically validate predictions from a specified class of scientific models against one relevant empirical observation to produce a score indicating agreement between the model and the data. 
Suites of such validation tests, collaboratively developed by a scientific community in common repositories, produce up-to-date statistical summaries of the state of the field. In this paper, we aim to detail this test-driven workflow and introduce it to the neuroscience community. As an initial example, we describe \textit{NeuronUnit}, a library that builds upon \textit{SciUnit} and integrates with several existing neuroinformatics resources to support validating single-neuron models using data gathered by neurophysiologists.

\tiny
 \keyFont{ \section{Keywords:} Neuroinformatics Simulation Electrophysiology Software Modeling Validation\todo{is this the right format for these?} } %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

% Section 1.
\section{Introduction}
Neuroscientists construct quantitative models to coherently explain experimental observations of neurons, circuits, brain regions and behavior. 
A model can be characterized by its \textit{scope}: the set of observable quantities that it can generate predictions about, and by its \textit{validity}: the extent to which its predictions agree with available experimental  observations of those quantities.

Today, scientists contribute a new model to the research community by submitting a paper containing a description of the model's structure and scope along with text and figures that demonstrate its validity and argue for its novelty.  
The scientists tasked with reviewing the paper are  responsible for evaluating these claims, discovering competing models and relevant data the paper did not adequately consider, and ensuring that goodness-of-fit was measured in a statistically sound manner, drawing on their knowledge of prior literature. Once a modeling paper is published, it is frozen and cited as-is in perpetuity.

Because publications are the sole vehicle for knowledge about model scope and validity, scientists must rely on an encyclopedic knowledge of the literature to answer \emph{summary questions} like the following:
\begin{itemize}
\item Which models are capable of predicting the quantities I am interested in?
\item What are the best practices for evaluating the goodness-of-fit between these models and  data?
\item How well do these models perform, as judged to these metrics, given currently available data?
\item What other quantities can and can't these models predict?
\item What observations have not yet been adequately explained by any available model?
\end{itemize}

In some fields, like neuroscience, where the number of relevant publications being generated every year is growing rapidly \citep{jinha_article_2010}, these questions can be difficult for even conscientious senior scientists to answer comprehensively. For new scientists, this represents a serious barrier to entry.

%This suggests that tools that help researchers answer questions like these could help readers and reviewers develop a more comprehensive understanding of the state-of-the-art and improve the quality of the research being published as a result: to renewed calls for tools that help researchers .

%These questions are particularly relevant for young scientists and  scientists aiming to enter a new research area to answer.

%understanding of the state-of-the-art from publications containing only a few pieces of data, and then only that data available at the time of publication.

%These demonstrations also quickly go out of date. Finally, there is no easy way to summarize and compare the claims being made in modeling papers in order to synthesize an understanding of a research area as a whole. 
%A strength of publications are their focused descriptions of new data and models. 
%A weakness, however, is that evaluating the scope and validity of models against known data is intractable using a body of publications alone.

Professional software developers face similar issues\todo{probably cite our in submission paper here??}. 
They must understand the scope of each component of a complex software project and verify that each component achieves desired input/output behavior. 
But software developers do not verify components by simply documenting a few interesting inputs and corresponding outputs and presenting them to other developers for one-time review leading to an archived document. 
Rather, they typically follow a \emph{test-driven development} methodology by creating a suite of executable \emph{unit tests} that serve to specify each component's scope and verify its implementation on an ongoing basis as it is being developed and modified \citep{beck2003}. 
Each test individually checks that a small portion of the program meets a single correctness criterion. For example, a unit test might verify that one function within the program correctly handles malformed inputs. 
Collectively, the test results serve as a summary of the project as it progresses through its development cycle. 
Developers can determine which features are unimplemented or buggy by examining the set of failed tests, and progress can be measured in terms of how many tests the program passes over time. 
This methodology is widely adopted in practice.

%publications would be analagous to user manuals, rather than rigorous justifications that a program meets a specific specification. Unit tests serve this purpose for software, so we argue that the validation of neuroscience models by means of executable tests would be useful to scientists as a supplement the publication system.

Test-driven methodologies have started to see success in neuroscience as well, in the form of \emph{modeling competitions}. 
During these competitions, competitors develop and parameterize models based on publicly-available training data and submit them to a central server. There, submitted model are provided hidden testing data which they must use to produce predictions. These predictions are validated using publicly available criteria to produce summaries of the relative merits of different models, just as a test suite summarizes the state of a software project. 
Such competitions continue to drive important advances and improve scientists' understanding of their fields. 
For example, the quantitative single neuron modeling competition (QSNMC) \citep{jolivet_quantitative_2008} investigates the complexity-accuracy tradeoff among reduced models of excitable membranes; 
the ``Hopfield'' challenge \citep{hopfield_what_2000} tested techniques for generating neuronal network form given its function; 
the Neural Prediction Challenge sought the best stimulus reconstructions, given neuronal activity (\url{http://neuralprediction.berkeley.edu}); the Diadem challenge is advancing the art of neurite reconstruction (\url{http://www.diademchallenge.org}); and 
examples from other areas of biomedical research abound (e.g. \url{http://www.the-dream-project.org}).
% \begin{figure}%{r}{0.6\textwidth}
% %\vspace{-38px}
% \includegraphics[scale=0.75]{sciunit_overview.pdf}
% \caption{Proposal overview. 
% \small{i) The \textit{SciUnit} framework can generate discipline-specific libraries for tests generation and model interfacing.  
% \textit{NeuronUnit} is proposed and described here.  Experimental data guides test development; model generation informs and is informed by these \textit{SciUnit} libraries.  
% Test suites for OSB and QSNMC (described in the text) are being built. 
% %\vspace{-10px}
% \label{fig:sciunit_overview}
% \end{figure}
% \leavevmode
% \vspace{-0px}
\begin{figure}
\vspace{-38px}
\centering
\includegraphics[scale=0.6]{diagram1.pdf}
\vspace{-55px}
\caption{NeuronUnit overview. NeuronUnit is set of testing tools built upon the discipline-agnostic SciUnit framework. 
NeuronUnit can in principle test arbitrary neurophysiology models using arbitrary data but we provide here an example using models described in NeuroML as part of the \textit{Open Source Brain Project} (OSB, \citep{gleeson_open_2012}, \url{http://www.opensourcebrain.org}), and single neuron electrophysiology data available as part of the NeuroElectro Project (Neuroelectro, \citep{tripathy_neuroelectro:_2012}, \url{http://neuroelectro.org}).}  
\label{fig:sciunit_overview}
\end{figure}
\leavevmode
\vspace{-0px}

Each of these examples has leveraged \emph{ad hoc} infrastructure to support model validation. 
While the specific criteria used to evaluate models can vary widely between modeling domains, the underlying methodology is common and could be implemented once. 
Recognizing this, and inspired by unit testing practices, we have developed a discipline-agnostic framework for developing \emph{model validation test suites} called \textit{SciUnit} (available from \url{http://sciunit.scidash.org}) \todo{cite sciunit paper again here?}. 
In this paper, we will begin by detailing \textit{SciUnit}, focusing on examples from single-neuron physiology (Sec. \ref{sec:sciunit}).

\textit{SciUnit} contains validation logic common across scientific disciplines, but a particular discipline, such as neurophysiology, might have more specialized logic associated with it that can be shared amongst its sub-disciplines (e.g. hippocampal neurophysiology). We anticipate a collaborative workflow where this common logic is developed in common repositories on a social coding service, here \textit{Github}. For neurophysiology, we have developed such a repository, called \textit{NeuronUnit}. This repository contains common testing logic as well as bridges to existing informatics infrastructure to make it easy to import existing data and models. In particular, we will show how models described using \text{NeuroML} and provided freely by the \textit{Open Source Brain Project} (OSB, \cite{gleeson_open_2012}, \url{http://www.opensourcebrain.org}) can be tested in a fully automated fashion using published data curated by the \textit{NeuroElectro Project} (NeuroElectro, \cite{tripathy_neuroelectro:_2012}, \url{http://neuroelectro.org}), leveraging facilities from the \textit{NeuroTools} library (\url{http://neuralensemble.org/NeuroTools}) to extract features from model outputs (Sec. \ref{sec:neuronunit}). Particular sub-disciplines, like hippocampal neurophysiologists, can use this common logic to select relevant models and tests, forming particular test suites in common repositories. The end result of this workflow is a table like the one shown in Figure \ref{fig:sciunit_overview}, where submitted models can be compared based on the scores they achieve on different selected tests. These tables serve as a summary of the status of a modeling community, just as the results from a suite of unit tests serve as a summary of the status of a software project, and help scientists answer the questions mentioned above more easily and accurately.

% Section 2.
\section{Validation Testing with \textit{SciUnit}}\label{sec:sciunit}
\subsection{Example: The Quantitative Single Neuron Modeling Competition}
\begin{figure}
\small
\begin{python}
class SpikeCountTest(sciunit.Test):
  """Tests spike counts produced in response to several current stimuli against observed means and standard deviations.

  goodness of fit metric: Computes p-values based on a chi-squared test statistic, and pools them using Fisher's method.
  parameters:
    inputs: list of numpy arrays containing input currents (nA)
    means, stds: lists of observed means and standard deviations, one per input
  """
  def __init__(self, inputs, means, stds):
    self.inputs, self.means, self.stds = inputs, means, stds

  required_capabilities = [SpikeCountFromCurrent]

  def _judge(self, model):
    inputs, means, stds = self.inputs, self.means, self.stds
    n = len(inputs)
    counts = numpy.empty((n,))
    for i in xrange(n):
      counts[i] = model.spike_count_from_current(inputs[i])
    chisquared = sum((counts-means)**2 / means) # An array of chi-squared values.  
    p = scipy.stats.chi2.cdf(chisquared,n-1) # An array of p-values.  
    pooled_p = sciunit.utils.fisherp(p_array) # A pooled p-value.  
    return sciunit.PValue(pooled_p, related_data={
      "inputs": inputs, "counts": counts, "obs_means": means, "obs_stds": stds
    })
\end{python}
\vspace{-5px}
\caption{An example single neuron spike count test class implemented using \textit{SciUnit}. Because this implementation contains logic common to many different systems, \textit{NeuronUnit} was developed to provide a simpler means to deliver it (see Sec. \ref{sec:neuronunit}).}
\label{fig:rate_test}
\vspace{-15px}
\end{figure}

%Simple executable \emph{validation tests} that compute agreement between a model prediction and an experimental observation.  
We first illustrate the form of a generic example \textit{SciUnit} test suite that could be used in neurophysiology. 
Suppose we have collected data from an experiment where current stimuli (measured in pA) are delivered to neurons in some brain region, while the somatic membrane potential of each stimulated cell (in mV) is recorded and stored.  
A model claiming to capture this cell type's membrane potential dynamics must be able to accurately predict a variety of features observed in these data.

One simple validation test would ask candidate models to predict the number of action potentials (a.k.a. spikes) generated in response to a stimulus (e.g. white noise), and compare these \emph{spike count} predictions to the distribution observed in repeated experimental trials using the same stimulus. 
For data of this type, goodness-of-fit can be measured by first calculating a p-value from a chi-squared statistic for each prediction and then combining these p-values using Fisher's method \citep{fisher_statistical_1925}.

Alongside this \emph{spike count test}, we might also specify a number of other tests capturing different features of the data  to produce a more comprehensive suite. 
For data of this sort, the QSNMC defined 17 other validation criteria in addition to one based on the overall spike count, capturing features like spike latencies (SL), mean subthreshold voltage (SV), interspike intervals (ISI) and interspike minima (ISM) that can be extracted from the data \citep{jolivet_quantitative_2008}. 
They then defined a combined metric favoring models that broadly succeeded at meeting these criteria, to produce an overall ranking. 
Such combined criteria are simply validation tests that invoke other tests to produce a result.

%In many cases, models require no modifications to take the new tests because the same type of model output is being requested.
\subsection{Implementing a Validation Test in \textit{SciUnit}}
Fig. \ref{fig:rate_test} shows how a scientist can implement spike count tests such as the one described above using \textit{SciUnit}. 
A \textit{SciUnit} validation test is an {instance} (i.e. an object) of a Python class implementing the \verbx{sciunit.Test} interface (cf. line 1). 
Here, we show a class \verbx{SpikeCountTest} taking three \emph{parameters} in its constructor (constructors are named \verbx{__init__} in Python, lines 9-10). 
The meaning of each parameter along with a description of the goodness-of-fit metric used by the test is documented on lines 4-7. 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} spike count test, we instantiate this class with particular experimental observations. 
For example, given observations from hippocampal CA1 cells (not shown), we can instantiate a test as follows\todo{use iPython style In[0] notation?}:
\begin{python}
  CA1_sc_test = SpikeCountTest(CA1_inputs, CA1_means, CA1_stds)
\end{python}
We emphasize the crucial distinction between the \textit{class} \verbx{SpikeCountTest}, which defines a \emph{parameterized family} of validation tests, and the particular \textit{instance} \verbx{CA1_sc_test}, which is an individual validation test because the necessary parameters, derived from data, have been provided. 
As we will describe below, we expect communities to build repositories of such families capturing the criteria used in their subfields of neuroscience so that test generation for a particular system of interest will often require simply instantiating a previously-developed family with particular experimental parameters and data. 
For single-neuron test families like \verbx{SpikeCountTest}, we have developed such a library, called \textit{NeuronUnit} (\url{http://github.com/scidash/neuronunit}) (Sec. \ref{sec:neuronunit}).

Classes that implement the \verbx{sciunit.Test} interface must contain a \verbx{_judge} method that receives a candidate \emph{model} as input and produces a \textit{score} as output. 
To specify the interface between the test and the model (that is, to specify an appropriate scope), the test author provides a list of \emph{capabilities} in the \verbx{required_capabilities} attribute, seen on line 12 of Fig. \ref{fig:rate_test}. 
Capabilities are simply collections of methods that a test will need to invoke in order to receive relevant data, and are analogous to \emph{interfaces} in e.g. Java (\url{http://docs.oracle.com/javase/tutorial/java/concepts/interface.html}). 
In Python, capabilities are written as classes with unimplemented members. 
The capability required by the test in Fig. \ref{fig:rate_test} is shown in Fig. \ref{fig:capability}. 
In \textit{SciUnit}, classes defining capabilities are tagged as such by inheriting from \verbx{sciunit.Capability}. The test in Figure \ref{fig:rate_test} uses this capability on line 19 to produce a spike count prediction for each input current.

The remainder of the \verbx{_judge} method implements the goodness-of-fit metric described above, returning an instance of \verbx{sciunit.PValue}, a subclass of \verbx{sciunit.Score} that is included with \textit{SciUnit}. In addition to the $p$-value itself, the returned score object also contains metadata, via the \verbx{related_data} parameter, for  scientists who may wish to examine the result in more detail later. 
In this case we save the input currents, the model outputs and the observed means and standard deviations (line 24).

\begin{figure}
\begin{python}
class SpikeCountFromCurrent(sciunit.Capability):
  def spike_count_from_current(self, input): 
    """Takes a numpy array containing current stimulus (in nA) and
    produces an integer spike count. Can be called multiple times."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\caption{An example capability specifying a single required method (used by the test in Figure \ref{fig:rate_test}).}
\label{fig:capability}
\vspace{-10px}
\end{figure}

\begin{figure}
\begin{python}
class TrainSpikeCountFromCurrent(sciunit.Capability):
  def train_with_currents(self, currents, counts):
    """Takes a list of numpy arrays containing current stimulus (in nA) and
    observed spike counts. Model parameters should be adjusted based on this
    training data."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\caption{Another capability specifying a training protocol (not used by the test in Figure \ref{fig:rate_test}).}
\label{fig:training}
\vspace{-15px}
\end{figure}

\subsection{Models}
Capabilities are \emph{implemented} by models. In \textit{SciUnit}, models are instances of Python classes that inherit from \verbx{sciunit.Model}. Like tests, the class itself represents a family of models, parameterized by the arguments of the constructor. A particular model is an instance of such a class.

Figure \ref{fig:simple_model} shows how to write a simple {family} of models, \verbx{LinearModel}, that implement the capability in Fig. \ref{fig:capability} as well as another capability shown in Fig. \ref{fig:training}, which we will discuss below. 
Models in this family generate a spike count by applying a linear transformation to the mean of the provided input current. The family is parameterized by the scale factor and the offset of the transformation, both scalars. 
To create a \emph{particular} linear model, a modeler can provide particular parameter values, just as with test families:
\begin{python}
CA1_linear_model_heuristic = LinearModel(3.0, 1.0)
\end{python}
Here, the parameters to the model were picked by the modeler heuristically, or based on externally-available knowledge. 
An alternative test design would add a training phase where these parameters were fit to data using the capability shown in Fig. \ref{fig:training}. 
This test could thus only be used for those models for which parameters can be adjusted without human involvement. 
Whether to build a training phase into the test protocol is a choice left to each test development community.

Fig. \ref{fig:rate_test} does not include a training phase. 
If training data is externally available, models that nevertheless do implement a training capability (like \verb|LinearModel|) can simply be trained explicitly by calling the capability method just like any other Python method:
\begin{python}
CA1_linear_model_fit = LinearModel()
CA1_linear_model_fit.train_with_currents(CA1_training_in, CA1_training_out)
\end{python}
\begin{figure}
\begin{python}
class LinearModel(sciunit.Model, SpikeCountFromCurrent, 
    TrainSpikeCountFromCurrent):
  def __init__(self, scale=None, offset=None): 
    self.scale, self.offset = scale, offset

  def spike_count_from_current(self, input):
    return int(self.scale*numpy.mean(input) + self.offset)

  def train_with_currents(self, currents, counts):
    means = [numpy.mean(c) for c in currents]
    [self.offset, self.scale] = numpy.polyfit(means, counts, deg=1)    
\end{python}
\caption{A model that returns a spike count by applying a linear transformation to the mean input current. The parameters can be provided manually or learned from data provided by a test or user (see text).}
\label{fig:simple_model}
\vspace{-15px}
\end{figure}

\subsection{Executing Tests} A test is executed against a model using the \verbx{judge} method:
\begin{python}
score = CA1_sc_test.judge(CA1_linear_model_heuristic)
\end{python}

This method proceeds by first checking that the provided model implements all required capabilities. It then calls the test's \verbx{_judge} method to produce a score. A reference to the test and model are added to the score for convenience (accessible via the \verbx{test} and \verbx{model} attributes, respectively), before it is returned.

\subsection{Test Suites and Score Matrices} A collection of tests intended to be run on the same model can be put together to form a test suite.
The following is a test suite that could be used for a simplified version of the QSNMC, as described above:  
\begin{python}
CA1_suite = sciunit.TestSuite([CA1_sc_test, CA1_sl_test, CA1_sv_test, CA1_isi_test, CA1_ism_test])
\end{python}
Like a single test, a test suite is capable of judging one or more models. The result is a score matrix much like the one diagramed in Fig. \ref{fig:sciunit_overview}.
\begin{python}
CA1_matrix = CA1_suite.judge([CA1_linear_model_heuristic, CA1_linear_model_fit])
\end{python}
A simple summary of the scores in a score matrix can be printed to the console or visualized by other tools in development.

% Section 3.
\section{Integrating With Neuroinformatics Tools and Infrastructure}\label{sec:neuronunit}
In the examples above, tests were parameterized by data explicitly, and models implemented capabilities directly. In many cases, however, both models and data can be accessed from specialized repositories in standardized formats. In this section, we will show how to leverage this existing infrastructure to make parameterizing tests with published data and generating \textit{SciUnit} models from models implemented in some other manner (even in a language other than Python) nearly automatic.

\subsection{\textit{NeuronUnit}: Neurophysiology-Specific Testing Tools}\label{sec:neuronunit_acitivities}
In particular, we will focus on neurophysiology in this paper as our initial case study, using machine-readable models written in \textit{NeuroML} available from OpenSourceBrain (\textit{OSB}), and machine-readable data from resources like \textit{NeuroElectro}. We will provide functions that  link these tools to \textit{SciUnit}, as a collection of relevant model and test families and associated capabilities, in a package called \textit{NeuronUnit}, collaboratively developed on Github (\url{http://github.com/scidash/neuronunit}) \todo{Must change name of repository on github to neuronunit (after I am sure grant has been reviewed)}. \todo{Sentence about how great github is for collaboratively doing things, particularly in science?} Scientists can use these tools to create test suites tailored to specific systems of interest (for example, scientists interested specifically in models of cells in the hippocampus would use \textit{NeuronUnit} to create a suite, \textit{HippocampusSuite}, parameterized by relevant data from \textit{NeuroElectro}).

We anticipate that other communities within neuroscience (and indeed, within science more broadly) will develop similar repositories that synergistically link \textit{SciUnit} to relevant ongoing standardization efforts and provide a common, collaboratively-developed vocabulary of test families, model families and capabilities relevant to their domain.

\subsection{Models from NeuroML}\label{sec:neuroml_models}
NeuroML is a standardized model description language for neuroscience \citep{gleeson_neuroml:_2010}. 
It permits many neurophysiological/neuroanatomical models to be described in a simulator-independent fashion, and executed across many popular simulators due to inter-conversion capabilities of the NeuroML API. 
Because NeuroML is an XML specification, model descriptions can be validated for correctness and queried for model properties and components, exposing potential capabilities. 
It is ideal for model sharing, curation, and for answering both \textit{what} and \textit{how} programatically.

NeuroConstruct (\cite{gleeson_neuroconstruct:_2007}, \url{http://www.neuroconstruct.org}) is a simulation manager that takes NeuroML models and hands off simulation to supported simulators. 
\textit{NeuronUnit} offers a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel}, instantiated with the path to a NeuroML model.  
Because NeuroML can describe such a wide range of models, \verbx{NeuroConstructModel} makes few assumptions about them: that each one is \verbx{TimeIntegrable}, and \verbx{HasMembranePotential}.  
It is subclassed to test \textit{specific} NeuroML models (Fig. \ref{fig:ca1_model}).

\textit{OSB} curates many models described in NeuroML. 
OSB-curated projects are converted from their native format into NeuroML, and run on major neural simulators such as Neuron (\cite{carnevale_neuron_2006}, \url{http://www.neuron.yale.edu/neuron}), GENESIS (\cite{bower_genesis_2007}, \url{http://genesis-sim.org}), NEST (\cite{gewaltig_nest_2007}, \url{http://www.nest-initiative.org}), and MOOSE (\cite{ray_moose_2008}, \url{http://moose.ncbs.res.in}). 
Concordance between model output (beginning with the NeuroML description) and reference output (from native simulator source files) is reported for each model. 
Thus, OSB is an excellent source of models that, in addition to being open source, are sufficiently described to enable validation. 
The hippocampal CA1 pyramidal cell is commonly modeled, and we implement one such model hosted on OSB (\url{http://opensourcebrain.org/projects/ca1pyramidalcell}) by simply declaring a \verbx{CA1PyramidalCellModel} class, inheriting from \verbx{NeuroConstructModel}.  
This basic implementation simply ``wraps'' the components of the existing model, with simulator interaction taken care of by \verbx{NeuroConstructModel} methods; 
thus, only the code shown in Fig. \ref{fig:ca1_model} is required. 
All OSB models, and indeed any NeuroML model, can be tested similarly.

Spanning a range of scales and original development environments, all OSB models are formally described using NeuroML, as are all model components and sub-components, such as cells, ion channels, calcium stores, etc. 
These models are regularly executed on OSB servers to ensure that their output remains consistent as they are updated. 
Therefore, OSB can confirm that they \textit{do} work, while linked journal articles, on-site wiki, and code inspection can establish \textit{how} they work. 
However, there is no mechanism for establishing \textit{how well} they work, i.e. how well the models accord with data. 
\textit{NeuronUnit} fills this gap by helping OSB (and the larger biology community) assess models using data-driven tests. 
\textit{SciUnit} can be applied similarly to other neuroscience (or biology) sub-disciplines using \textit{NeuronUnit} analogues written by the corresponding communities.

\begin{figure}
\begin{python}
class CA1PyramidalCellModel(NeuroConstructModel):
	"""CA1 Pyramidal Cell model from Open Source Brain."""
	def __init__(self,**kwargs):
		project_path = neuroconstruct.get_path("hippocampus","CA1_pyramidal_neuron","CA1PyramidalCell","neuroConstruct")
		models.NeuroConstructModel.__init__(self,project_path,**kwargs)
\end{python}
\vspace{-5px}
\caption{A model class corresponding to a CA1 Pyramidal Cell model from Open Source Brain}
\label{fig:ca1_model}
\vspace{-15px}
\end{figure}

\subsection{Capabilities from NeuroTools}
NeuroTools (\url{http://neuralensemble.org/NeuroTools}) is a Python library supporting tasks associated with analysis of neural data (or model output), such as membrane potential time series, spike trains, etc. 
It is an open source and actively developed project, containing reliable algorithms on which to base neurophysiology tests.

We use NeuroTools to implement \textit{SciUnit} capabilities in \textit{NeuronUnit}. 
For example, an \verbx{AnalogSignal} object (e.g. a membrane potential time series) from NeuroTools has a threshold detection method that returns a NeuroTools \verbx{SpikeTrain} object. 
A \textit{NeuronUnit} \verbx{HasSpikeTrain} Capability requires that the method \verbx{getSpikeTrain} be implemented. 
\verbx{NeuroConstructModel} does so by placing the object method \verbx{AnalogSignal.threshold_detection} inside \verbx{getSpikeTrain}. 
Many such NeuroTools objects are similarly exchanged between \verbx{NeuroConstructModel} methods. 
This simplifies test writing, since basic model output properties are obtained trivially using NeuroTools object methods, and these NeuroTools objects are easily extracted from model output using candidate models subclassing \verbx{NeuroConstructModel}.

\subsection{Reference Data for Tests from NeuroElectro}
Answering \textit{how well} requires validation testing against data. 
The NeuroElectro project (\url{http://neuroelectro.org}) is an effort to curate all published single cell neurophysiology data \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are reported for 93 cell types, spanning $>$ 2000 single pieces of published data extracted from article tables. 
We have made it easy to construct \textit{NeuronUnit} tests using the NeuroElectro API to get reference data. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The former is illustrated in Figure \ref{fig:neuronunit_example}. 
Associated statistics of that data (e.g mean, standard error, and sample size) are attached and enable judgement of model output according to a chosen scoring mechanism. 
While NeuroElectro alone cannot judge all model aspects, it can serve to validate basic features of many neurophysiology models, such as resting membrane potential, action potential width, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key component for \textit{NeuronUnit} test constuction.

\begin{figure}%[18]{r}{1\textwidth}
\includegraphics[scale=0.65]{table.png}
\caption{A record matrix for SciUnit tests}
\label{fig:scidash_matrix}
\vspace{-15px}
\end{figure}
\vspace{-10px}
\leavevmode

\section{SciDash: A Lightweight Portal and Visualization Tool}\label{sec:scidash}
% Section 4.
TODO: Short summary of how SciDash works

\section{Discussion}\label{sec:discussion}

\subsection{A Complete Pipeline}
Although the tools described in Sec. \ref{sec:neuronunit_acitivities} do not exhaust the possible sources of models, capabilities, and test data, they provide an immediate point of entry into the neurophysiology community and a powerful demonstration of our proposal. 
In the \textit{NeuronUnit} repository (\url{http://github.com/scidash/neuronunit}) is a runnable script (\textit{examples.py}) demonstrating a complete testing pipeline. 
It (1) selects an OSB model; 
(2) simulates it using NeuroConstruct; 
(3) tests the widths of the resulting action potentials, extracted and computed using NeuroTools, against NeuroElectro data downloaded on-the-fly, using a \textit{NeuronUnit} test class called \verbx{SpikeWidthTestDynamic}; 
and (4) computes and prints a test score (Figure~\ref{fig:neuronunit_example}).

\begin{figure}
\begin{python}
# Interface with neuroelectro.org to search for spike widths of CA1 Pyramidal cells. 
reference_data = NeuroElectroSummary(neuron={'id':85}, ephysprop={'id':23})
reference_data.get_values()  # Get summary data for the above. 
model = CA1PyramidalCellModel(population_name="CG_CML_0") # Initialize the model with some parameters.
test = SpikeWidthTestDynamic( # Initialize the test.    
	reference_data = {'mean':reference_data.mean, 'std':reference_data.std}, # Summary statistics from the reference data
	model_args = {'current':40.0}, # Somatic current injection in pA.  
	comparator = ZComparator), # A comparison class that implements a Z-score.  
result = sciunit.judge(test,model) # (1) Check capabilities, (2) take the test, (3) generate a score and validate it, (4) bind the score to model/test combination. 
result.summarize() # Summarize the result.  
\end{python}
\caption{Working example of testing in \textit{NeuronUnit}}
\label{fig:neuronunit_example}
\vspace{-15px}
\end{figure}

\subsection{Creating New Models, Capabilities, and Tests}
\textit{NeuronUnit} provides base classes to enable rapid generation of models, capabilities, and tests for neurophysiology data. 
However these objects can also be created from scratch, requiring only adherence to the \textit{SciUnit} interface. 
For example, a Model could implement an \verbx{integrate} capability method by wrapping execution of a MATLAB script and a \verbx{get_spikes} capability method by parsing a .csv file on disk; 
a Test could be initialized using empirical spike rates collected in the lab.  
While this does not meet our idealized vision of development and testing, in practice this may be a common scenario. 
Adoption of \textit{SciUnit} into traditional, custom workflows would be reflected in the Methods sections of articles.

\subsection{Participation from Modeling Communities}
Modelers may not want to expose model capabilities, a requirement of test-taking.  
We anticipate four solutions: \textbf{First}, interfacing a model to \textit{SciUnit} requires only implementing selected model capabilities. 
Often this means identifying native model procedures that satisfy a capability, and wrapping their functionality. 
This can require as little as one line of code. 
Importantly, the modeler is not required to expose or rewrite any model flow control. 
\textbf{Second}, we support multiple environments automatically by using NeuroML \citep{gleeson_neuroml:_2010}, and other simulator-independent model descriptions are possible for other domains. 
Automated generation of NeuroML from native model source code is in development (Gleeson, personal communication); for the popular NEURON simulator (\cite{carnevale_neuron_2006}), this functionality is already mature and in use. 
This minimizes modeler effort for a large (at least 1000, \url{http://www.neuron.yale.edu/neuron/node/69}) and growing number of neuronal models. 
\textbf{Third}, modelers have an incentive to demonstrate publicly their models' validity. 
Participation in public modeling competitions demonstrates this incentive. 
\textbf{Fourth}, modelers have an incentive to use \textit{SciUnit} during development (see TDD, above) to ensure that ongoing development preserves correspondence between model and data. 
A popular test suite can represent a ``gold standard'' by which progress during development is judged.

\subsection{Participation from Experimental Communities}
Experimentalists may not want to write tests derived from their data.  We anticipate four solutions: \textbf{First}, tests require no special data formatting; only a list of required  capabilities (for selecting eligible models), optional metadata (as run-time arguments), and a statistical data summary (for scoring tests) are required.  A unit test is focused and does not require arbitrary computations on data.  For example, suppose intracellular current injection evokes 100 action potentials, the width of which is of interest.  Writing the test consists of selecting \verbx{ReceivesCurrent} and \verbx{ProducesActionPotentialShape} capabilities (one line of code each), computing the mean and variance of action potential widths (one line of code), specifying current injection parameters, e.g. amplitude and duration (two lines of code), and selecting a scoring mechanism from \verbx{sciunit.scores}, e.g. (colloquially) ``Must be $<$ 1 standard deviation of the mean'' (one line of code).  This example can be found in \verbx{NeuronUnit.tests.SpikeWidthTest}; heavy-lifting is done by the interface.
\textbf{Second}, data-sharing is becoming accepted, and test-writing can be distributed across scientists, including non-experimentalists with other aims such as analysis or modeling. 
\textbf{Third}, many tests can be automatically generated using the NeuroElectro API, and the continued emergence of such data-aggregation initiatives will expand these possibilities. 
\textbf{Fourth}, an incentive to write tests for one's data exists: the ability to identify models that give the data clear context and impact.

\subsection{Diversity of Levels and Kinds of Models and Data}
The diversity of topics in biology is vast. 
\textbf{First}, we address this by providing an interface allowing modelers to express specific capabilities.  
This capability set determines the range of eligible tests.  Scale hierarchies are embedded in capability inheritance.  =
For example, \verbx{HasActionPotentials} inherits from \verbx{HasNeurons}, and \verbx{HodgkinHuxley} inherits from \verbx{VoltageGated}. 
Thus, incompatibility of a test-requiring-action-potentials for a model-lacking-neurons is known without explicit tagging. 
\textbf{Second}, NeuroML naturally addresses diversity of scales because it is organized hierarchically, in ``levels.''  
Models can be sub- or supersets of other models; similarly for SBML (\cite{hucka_systems_2003}, \url{http://sbml.org}) a general systems biology markup language. 
\textbf{Third}, cross-level testing can use ``Representional Similarity Analysis'' (RSA) \citep{kriegeskorte_representational_2008}, requiring only that a model respond to defined inputs (e.g. stimuli). 
A ``similarity matrix'' for input responses defines a unique model signature, and can serve as intermediate test output. 
Goodness-of-fit between similarity matrices for model and experiment determines test scores; 
these matrices are independent of model scale because their size depends only on test inputs, not system detail.

\subsection{Arbitrary Scoring Criteria for Tests}
A test first assesses goodness-of-fit, and applies a normalization (e.g. pass/fail, 0.0-1.0) to generate a score. 
Arbitrary choices at both stages may benefit some models over others.  \textbf{First}, however, rank-ordering is constant across many goodness-of-fit metrics, meaning that choice of metric will rarely cause an otherwise passing model to fail and vice versa. 
For example, given a data mean and variance, ordering model output by Z-score or p-value will yield the same relative ranking of models. 
Indeed, rank ordering of models may prove more valuable than test scores themselves. 
\textbf{Second}, suite repositories should be open, so tests can be cloned and new statistical choices implemented. 
Statistics as applied to neuroscience have been notoriously ``fast and loose''; 
identification and correction of flawed methodology is becoming increasingly common (\cite{button_power_2013,kriegeskorte_circular_2009,galbraith_study_2010}, \url{http://prefrontal.org/files/posters/Bennett-Salmon-2009.jpg}), and is accelerated by an open framework. 
The validation criteria is made explicit in the specification of a test, so modelers need not guess which criteria are being used to validate or invalidate their model. 
Validation criteria are subject to debate (indeed, the QSNMC criteria changed between 2007 and 2008 due to such debates), and neuroscientists who wish to promote different criteria need only derive alternative tests. 
A community consensus will slowly emerge in the form a commonly-accepted test suite. 
This community can judge which test version is most appropriate, i.e. what a model \textit{should} do -- this process documented via standard moderation techiniques used on GitHub.  This process can be made even more visible by storing the output of test runs using Sage or IPython worksheets, or through optional database backends such as in the Sumatra lab notebook for simulation and analysis (\cite{sumatra_davison_2012}, \url{http://neuralensemble.org/sumatra}).

\subsection{Reliability of Data Underlying Tests}
Unreliable data can undermine model validation. 
\textbf{First}, the community must evaluate experimental design and methods, discounting data produced using questionable techniques. 
GitHub supports community moderation, permitting users to comment on tests, indicating their concerns. 
Suite repository popularity can reflect consensus. 
Experimental metadata also constrains a test's relevance, so test writers should select data with metadata appropriate to the system being modeled, and attach the metadata to resulting test scores. 
Metadata can also be expressed as Capabilities, e.g. \verbx{At37Degrees} or \verbx{Calcium3mM}; and tests can require that models express them. 
Such capabilities require no implementation, so the model definition must only inherit them. 
\textbf{Second}, models cannot perfectly reproduce data that is itself a random draw from a ``true'' distribution. 
Uncertainty in data must be made explicit, by asking how well a data set validates its own experimental replications \citep{kriegeskorte_representational_2008}. 
The degree of such ``self-validation'' represents the upper limit of what a model can be expected to achieve, and should represent a ``perfect'' score.

\subsection{Occam's Razor}
All things being equal, simpler models are better. 
Model complexity has many definitions, so \textit{SciDash} will report several complexity metrics \citep{mccabe_complexity_1976}, including: 
1) model length; 2) memory use; 3) CPU load; 4) \# of capabilities. 
Future \textit{SciUnit} tools will report the model validity vs complexity tradeoff in tabular form (e.g. Table \ref{fig:scidash_matrix}), and in a scatter plot, with the ``best'' models being in the high validity / low complexity corner of the plot. 
The set of models which \textit{dominate} all others, i.e. that have the highest validity for a given complexity, can be represented as a ``frontier'' in such a scatter plot, a visualization familiar from the symbolic regression package Eureqa \citep{schmidt_distilling_2009}.

\subsection{Expansion Into Other Areas of Biology}
After covering neurophysiology, we would like \textit{SciUnit} to be applied across neuroscience and in other biological sciences. 
The framework is discipline-agnostic, so community participation and model description are the only obstacles. 
Community participation begins with enumerating the capabilities relevant to a sub-discipline, and then writing tests. 
Model description can expand within NeuroML (which already covers multiple levels within neuroscience) and tools for capability implementation can incorporate libraries for neuroimaging (NiBabel, \url{http://nipy.org/nibabel}), neuroanatomy (NeuroHDF, \url{http://neurohdf.readthedocs.org}) and other sub-disciplines. 
SBML will enable expansion beyond neuroscience, facilitated by parallel efforts among NeuroML developers to interface with it (Crook, unpublished). 
One intriguing possiblity is applying \textit{SciUnit} to the OpenWorm project (\url{http://www.openworm.org}), which through open, collaborative development seeks to model the entire organism C. elegans.

\section*{Disclosure/Conflict-of-Interest Statement}
%All relationships financial, commercial or otherwise that might be perceived by the academic community as representing a potential conflict of interest must be described. If no such relationship exists, authors will be asked to declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Acknowledgements}
We thank Sharon Crook, Jonathan Aldrich, Shreejoy Tripathy, and Padraig Gleeson for their support of this project.  
The work was supported in part by grant R01MH081905 from the National Institute of Mental Health. 
The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health.

%\bibliographystyle{frontiersinSCNS&ENG} % for Science and Engineering articles
\bibliographystyle{frontiersinSCNS&ENG} % for Medicine articles
\bibliography{references,references_crook,urls,references_ja}

%\bibliographystyle{unsrt}
%\bibliography{references,references_crook,urls,references_ja}
%\listoftodos
\end{document}